2.6: Inferential Statistics I

### 2019-10-03 content/class/12-class.Rmd
There's no practice for class 12, shorten slides 12

- `broom`: for tidy regression outputs, summary statistics, and adding $\hat{Y_i}$ and $\hat{u_i}$ into the dataframe
- `huxtable`: to present regression output in a table with `huxreg()`
- `lmtest`: for testing for heteroskedasticity in errors with `bptest()`
- `car`: for testing for outliers with `outlierTest()`
- `estimatr`: for calculating robust standard errors with `lm_robust()`

@@

5. Intepret the confidence interval in the context of the problem

@@

## Hypothesis Testing

>>>>

- `infer`: for simulation for statistical inference

@@

5. **Intepret the confidence interval in the context of the problem**

@@

### 2019-10-03 static/slides/12-slides.Rmd
There's no practice for class 12, shorten slides 12

# Distributions of the OLS Estimators

@@

# Inferential Statistics and Sampling Distributions

@@

- We almost never can directly study the population, so we *model* it with our samples 
]

@@

# Inference in Econometrics: The Big Picture

@@

1. A .onfire[null hypothesis], $H_0$**

2. An .onfire[alternative hypothesis], $H_a$**

@@

2. .shoutcaps[Type II error (false negative)]: failing to reject $H_0$ when it is in fact false
    - Believing we found nothing when there was truly a relationship to find
]

@@

.content-box-blue[
.blue[**Let's Check**:]
For each of the following scenarios, identify the Type I error, Type II error, and which error is of greater consequence:
]

--
1. $H_0:$ the victim of an automobile accident is alive when he arrives at the Emergency Room

2. $H_0$: a rock climber's equipment is safe

3. $H_0$: a woman is not pregnant

4. $H_0$: a highway project will cost no more than $10 million

5. $H_0$: an experimental cancer drug has a cure rate of at least 75%

@@

- The .shout[confidence level] is defined as .shout[`\\((1-\alpha)\\)`]

- We often specify in advance an $\alpha$-level (0.10, 0.05, 0.01) with associated confidence level (90%, 95%, 99%)

@@

.regtable[
```{r, echo=FALSE, results="asis"}
tribble(
  ~"", ~"", ~"Null is True", ~"Null is False",
  "Judgment", "Reject Null", "Type I Error", "Correct",
  "Judgment", "Reject Null", "alpha", "1-beta",
  "Judgment", "Don't Reject Null", "Correct", "Type II Error",
  "Judgment", "Don't Reject Null", "1-alpha", "beta"
  ) %>%
  knitr::kable(.,format="html") %>%
  kable_styling(full_width = F) %>%
  add_header_above(c(" " = 2, "Truth" = 2)) %>%
  column_spec(c(1,2), bold=T) %>%
  collapse_rows(columns = c(1,2), latex_hline = "major", valign = "middle")
```
]

@@

- The statistical .shout[power of the test] is $1-\beta$, the probability of correctly rejecting $H_0$ when $H_0$ is in fact false (e.g. not convicting an innocent person)

@@

$$p(|H_0)$$

- If $p<\alpha$, the results are "**statistically significant**"

@@

- After running our test, we need to make a *decision* between the competing hypotheses
- Compare $p$-value with *pre-determined* $\alpha$ (commonly, $\alpha=0.05$, 95% confidence level)
    - If $p<\alpha$: **statistically significant** evidence sufficient to *reject* $H_0$ in favor of $H_a$
    - If $p \geq \alpha$: *insufficient* evidence to reject $H_0$
        - Note this does **not** mean $H_0$ is true! We merely have *failed* to *reject* $H_0$

@@

- Modern philosophy of science is largely based off of hypothesis testing and .whisper[falsifiability], which form the "Scientific Method"<sup>1</sup>
- For something to be "scientific", it must be .whisper[falsifiable], or at least .whisper[testable]
- Hypotheses can be *corroborated* with evidence, but always *tentative* until falsified by data in suggesting an alternative hypothesis
- .onfire["All swans are white"] is a hypothesis rejected upon discovery of a single black swan
]

@@

# Hypothesis Testing I

@@

# Hypothesis Testing II

- Rigorous course on statistics (**ECMG 212** or **MATH 112**) will spend weeks going through different types of tests:
    - Sample mean; difference of means
    - Proportion; difference of proportions
    - Z-test vs t-test
    - 1 sample vs. 2 samples
    - $\chi^2$ test

- Fortunately, some statisticians realized ["**there is only one test**"](https://allendowney.blogspot.com/2011/05/there-is-only-one-test.html) and coded it that way

@@

1. Calculate a statistic, $\delta$<sup>1</sup>, from a sample of data
2. Simulate a world where $\delta$ is null
3. Examine $\delta$ in the null world
4. Calculate the probability that $\delta$ could exist in the null world
5. Decide if $\delta$ is statistically significant

.footnote[<sup>1</sup> $\delta$ can stand in for any test-statistic in any hypothesis test! For our purposes, $\delta$ is the slope of our regression sample, $\hat{\beta}_1$.]

@@

# Test Statistic

- .whisper[Test statistic]: measure of how far what we observed in our sample is from what we would expect if the null hypothesis was true
    - Calculate from a distribution
- .whisper[Rejection region]: if the test statistic reaches a .onfire["critical value"], then we reject the null hypothesis

# Statistical Significance and Economic Significance

## Is a finding *significant* ?

## Is a finding *meaningful* ?

# How Big is "Big"?

.center[
![:scale 40%](https://www.dropbox.com/s/qtd2uwvr52jvkcu/drwhocontext.jpg?raw=1)
]

# Imagine a Null World

.center[
![:scale 70%](https://www.dropbox.com/s/x7b5azt5algjodg/parallelworld.jpeg?raw=1)
]

# Imagine a Null World

# Our Sample

```{r, echo=F}
tidy(school_reg)
```
]

# Another Sample

```{r, echo = F}
library(infer)
CASchool %>%
  specify(testscr ~ str) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1, type = "permute") %>%
  lm(testscr ~ str, data = .) %>%
  tidy()
```
]

# 1000 Parallel Universes

.left-code[

```{r infer-example, echo = T, eval = F}
library(infer)
simulations <- CASchool %>%
  specify(testscr ~ str) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")

simulations %>%
  rename(sample = replicate,
         slope = stat)
```
]

.right-plot[
```{r ref.label="infer-example", echo = F}
```

# 1000 Parallel Universes

```{r echo = F}
plot <- simulations %>%
  visualize()+
  theme_minimal()+
  labs(x = "Slope",
       y = "Samples")

plot
```

# 1000 Parallel Universes

```{r}
obs_slope <- school_reg_tidy %>% 
  filter(term=="str") %>%
  pull(estimate) # from dplyr

plot2 <- simulations %>%
  visualize(obs_stat = obs_slope)

plot2+
  shade_p_value(obs_stat = obs_slope, 
                direction = "two_sided")+
  theme_minimal()+
  labs(x = "Slope",
       y = "Samples")
```

# 1000 Parallel Universes

```{r}
# get confidence interval values

school_reg_CI<- school_reg %>%
  tidy(., conf.int=T) %>%
  filter(term=="str") %>%
  select(conf.low, conf.high)
```

```{r}
plot2+
  shade_confidence_interval(endpoints = school_reg_CI)+
  theme_minimal()+
  labs(x = "Slope",
       y = "Samples")
```

# p-Value

```{r}
simulations %>%
  get_p_value(obs_stat = obs_slope, direction = "both")
```

> There's a 0% chance of seeing a slope at least as large as $-2.28$ in the null world where there is no relationship.

# What R Does

```{r}
tidy(school_reg)
```

- R calculates the $p$-value the "old-fashioned way", using a theoretical distribution instead of simulation
    - Generates a test statistic (`statistic`) and compares it to a critical value of the test statistic<sup>1</sup>
$$\text{test statistic} = \frac{\text{estimate} - \text{null hypothesis}}{\text{standard error of estimate}}$$

.footnote[<sup>1</sup> Rule of thumb: test statistic of 2 or larger is generally significant at the 95% level. This comes directly from the empirical rule that 95% of observations occur within 2 standard deviations of the mean! See the **class info** page for more.]

# What R Does II

```{r}
tidy(school_reg)
```

$$t = \frac{-2.28-0}{0.48} = -4.75$$
- Intepretation like $Z$-scores: number of standard deviations away from the center<sup>1</sup>
- $p$-value is the probability of getting a test statistic at least as large (in magnitude) as this one if the null hypothesis were true<sup>2</sup>

.footnote[<sup>1</sup> The "center" here is the null hypothesis value i.e. a slope of 0.
<sup>2</sup> Think of our simulated null world(s), the center of the sampling distribution was a slope of 0.]

# Confidence Intervals

```{r}
tidy(school_reg, conf.int=TRUE)
```

- We can also generate a .shout[confidence interval] for $\hat{\beta}_1$ that contain the true population slope ($\beta_1$) with a specified probability
    - Note confidence level $=1-\alpha$; commonly 95% ($\alpha=0.05$)
- .whisper[We are 95% confident that the true slope is between $-3.22$ and $-1.34$.]
    - Note if the confidence interval contains 0, *cannot* reject the null hypothesis of $\beta_1=0$! 
- See **class info** page for more on confidence intervals and how they are constructed

```{r, results="asis"}
equatiomatic::extract_eq(school_reg, use_coefs = TRUE,coef_digits = 2,fix_signs = TRUE)
```

# Testing Reg Output: huxtable

.font80[
.regtable[
```{r, echo=F}
library(huxtable)

hr <- huxreg("Model 1" = school_reg, # can name the columns in quotes
       coefs = c("Intercept" = "(Intercept)", # can rename the coefficients
                 "STR" = "str"),
       statistics = c("N" = "nobs", # statistics to show for regression, taken from glance()
                      "R Squared" = "r.squared",
                      "Adj. R Squared" = "adj.r.squared",
                      "SER" = "sigma",
                      "F" = "statistic"),
       number_format = 2, # number of decimals to round to
       bold_signif = 0.05) # can bold coeffs by significance level

hr
```
]
]

>>>>

# Recall: Distributions of the OLS Estimators

@@

# Recall: Inferential Statistics and Sampling Distributions

@@

@@

# Recall: Inference in Econometrics: The Big Picture

@@

1. A .onfire[null hypothesis, `\\(H_0\\)`]

2. An .onfire[alternative hypothesis, `\\(H_a\\)`]

@@

2. .shout[Type II error (false negative)]: failing to reject $H_0$ when it is in fact false
    - Believing we found nothing when there was truly a relationship to find
]

@@

.center[
![](https://www.dropbox.com/s/wf6i1462dadmvrw/errorspregnant.png?raw=1)
]

@@

- The .shout[confidence level] is defined as .shout[`\\((1-\alpha)\\)`]
  - Specify *in advance* an $\alpha$-level (0.10, 0.05, 0.01) with associated confidence level (90%, 95%, 99%)

@@

.regtable[
```{r, echo=FALSE, results="asis"}
tribble(
  ~"", ~"", ~"Null is True", ~"Null is False",
  "Judgment", "Reject Null", "TYPE I ERROR", "CORRECT",
  "Judgment", "Reject Null", "(alpha)", "(1-beta)",
  "Judgment", "Don't Reject Null", "CORRECT", "TYPE II ERROR",
  "Judgment", "Don't Reject Null", "(1-alpha)", "(beta)"
  ) %>%
  knitr::kable(.,format="html") %>%
  kable_styling(full_width = F) %>%
  add_header_above(c(" " = 2, "Truth" = 2)) %>%
  column_spec(c(1,2), bold=T) %>%
  collapse_rows(columns = c(1,2), latex_hline = "major", valign = "middle")
```
]

@@

- The statistical .shout[power of the test] is $(1-\beta)$: the probability of correctly rejecting $H_0$ when $H_0$ is in fact false (e.g. not convicting an innocent person)

@@

$$p(\delta_i \geq \delta^*|H_0 \text{ is true})$$
    - where $\delta$ represents some test statistic
    - $\delta_i$ is the test statistic we observe in our sample
    - $\delta^*$ is some **critical value** of the test statistic
    - More on this in a bit

@@

- After running our test, we need to make a *decision* between the competing hypotheses

- Compare $p$-value with *pre-determined* $\alpha$ (commonly, $\alpha=0.05$, 95% confidence level)
    - If $p<\alpha$: **statistically significant** evidence sufficient to *reject* $H_0$ in favor of $H_a$
    - If $p \geq \alpha$: *insufficient* evidence to reject $H_0$
        - Note this does **not** mean $H_0$ is true! We merely have *failed* to *reject* $H_0$

@@

- Modern philosophy of science is largely based off of hypothesis testing and .onfire[falsifiability], which form the "Scientific Method"<sup>1</sup>

- For something to be "scientific", it must be .whisper[falsifiable], or at least .onfire[testable]

- Hypotheses can be *corroborated* with evidence, but always *tentative* until falsified by data in suggesting an alternative hypothesis

> **"All swans are white"** is a hypothesis rejected upon discovery of a single black swan 
]

@@

# Hypothesis Testing and p-Values

@@

# Hypothesis Testing: Which Test? I

- Rigorous course on statistics ([**ECMG 212**](http://ryansafner.com/courses/ecmg212) or **MATH 112**) will spend weeks going through different types of tests:
    - Sample mean; difference of means
    - Proportion; difference of proportions
    - Z-test vs t-test
    - 1 sample vs. 2 samples
    - $\chi^2$ test

- See today's [class notes](/class/12-class) page for more

@@

# Hypothesis Testing: Which Test? II

.center[
![:scale 80%](https://www.dropbox.com/s/3f33cnn595x2nsj/hypothesistestflowchart.png?raw=1)
]

# There is Only One Test

- Fortunately, some statisticians realized ["**there is only one test**"](https://allendowney.blogspot.com/2011/05/there-is-only-one-test.html) and some clever folks built a nice `R` package called `infer`

1. Calculate a statistic, $\delta$<sup>1</sup>, from a sample of data

2. Simulate a world where $\delta$ is null

3. Examine $\delta$ in the null world

4. Calculate the probability that $\delta$ could exist in the null world

5. Decide if $\delta$ is statistically significant

.footnote[<sup>1</sup> `\\(\delta\\)` can stand in for any test-statistic in any hypothesis test! For our purposes, `\\(\delta\\)` is the slope of our regression sample, `\\(\hat{\beta}_1\\)`.]

@@
